A Two-Infall Star Formation History with Outflows and Radial Migration

### 2025-02-21 src/tex/ms.tex

The dilution effect discussed in \ref{sec:yields} is clearly seen in the multi-zone model results. Figure \ref{fig:abundance-evolution} shows stellar age--abundance relations for four model outputs. Column (a) plots the output of our fiducial model, with $y/Z_\odot=1$, $\sigma_{\rm RM8}=2.7\,{\rm kpc}$, and $Z_{\rm CGM}=0$. The fiducial model shows two major discrepancies with the \citet{leung_variational_2023} age--abundance relation: a major dilution of $\sim0.5$ dex at a lookback time of $\sim9\,{\rm Gyr}$ near where the data show a maximum in [O/H], and non-zero abundance evolution at late times where the data show very little abundance evolution. We also note the present-day metallicity is still sub-Solar in the fiducial model, but this can be corrected by decreasing $\eta$. A higher yield set (column b) mitigates both of these issues by shortening the time it takes the ISM metallicity to rebound post-$t_{\rm max}$, producing a much flatter abundance curve at late times. However, a consequence of these higher yields is a poor fit to the stellar [O/Fe] distribution, as discussed in Section \ref{sec:abundance-distributions}.

@@

On the other hand, the model predictions for the MDF in the $8-10\,{\rm Gyr}$ age bin are strikingly similar to the APOGEE MDF. The two-infall model predicts a bimodal distribution: the metal-rich component comprises stars formed during the last Gyr of the first infall epoch, and the metal-poor component is stars formed during the first Gyr of the second infall epoch (post-dilution). The data show this same bimodal structure in all but the outermost radial bin, with peaks at ${\rm [Fe/H]}\approx-0.3$ and $+0.3$. While the location of the model peaks depends on $R_{\rm gal}$, in the data they are quite consistent from the innermost to the outermost bin.

>>>>

The dilution effect discussed in \ref{sec:yields} is clearly seen in the multi-zone model results. Figure \ref{fig:abundance-evolution} shows stellar age--abundance relations for four model outputs. Column (a) plots the output of our fiducial model, with $y/Z_\odot=1$, $\sigma_{\rm RM8}=2.7\,{\rm kpc}$, and $Z_{\rm CGM}=0$. The fiducial model shows two major discrepancies with the \citet{leung_variational_2023} age--abundance relation: a major dilution of $\sim0.5$ dex at a lookback time of $\sim9\,{\rm Gyr}$ near where the data show a maximum in [O/H], and non-zero abundance evolution at late times where the data show very little abundance evolution. The evolution of [Fe/H] is similar, but the approach to Solar abundance is slower because of the additional delay imposed on Fe production from SNe Ia. A higher yield set (column b) mitigates both of these issues by shortening the time it takes the ISM metallicity to rebound post-$t_{\rm max}$, producing a much flatter abundance curve at late times. However, model (b) produces a poorer fit to the age--[O/Fe] relation; the decline in [O/Fe] over the thin disk epoch is much steeper than the data.

@@

On the other hand, the model predictions for the MDF in the $8-10\,{\rm Gyr}$ age bin are strikingly similar to the APOGEE MDF. The two-infall model predicts a bimodal distribution: the metal-rich component comprises stars formed during the last Gyr of the first infall epoch, and the metal-poor component is stars formed during the first Gyr of the second infall epoch (post-dilution). The data show this same bimodal structure in all but the outermost radial bin, with peaks at ${\rm [Fe/H]}\approx-0.3$ and $+0.3$. While the location of the model peaks depends on $R_{\rm gal}$, in the data they are quite consistent from the innermost to the outermost bin. \todo{I actually think the two-infall model gets this right for the wrong reasons. In the data, the left peak is the high-$\alpha$ sequence and the right is the high-metallicity end of the low-$\alpha$ sequence. In the model, the {\it right} peak is produced during the first infall (i.e., thick disk) and the {\it left} is immediately after the second infall (i.e., low-metallicity end of thin disk).}
